TAPIR: Threaded Application Programming on Intermittent Resources

### 2017-05-11 paper.tex

\usepackage{balance}
\usepackage{comment}
\usepackage{listings}
\usepackage{url}
\usepackage{xspace}
\usepackage[export]{adjustbox}
\usepackage{wrapfig}
\newcommand{\chain}{Chain\xspace}
\numberofauthors{3}

@@

%\section{Implementation}
%Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
%Mutexes (Simple, meant for pseudo channels) - Emily\\
%Interrupts (mucked with IE flags) - Neil\\
%% Interrupt prologue
%% Interrupt setup

\section{Evaluation}
To test our multithreading library and mutex implementation, we instrumented
several existing Chain applications that do not share any tasks or channels to
run concurrently. The programs were run on either an msp430fr5969 launchpad or
on a Capybara sensing node, a custom intermittent development board with an
msp430fr5949 MCU. Reboot tests were carried out strictly on the Capybara board
because was simpler to guarantee the isolation of the MCU from the programming
fet than on the launchpad. We did not seekt to fully verify our implementation
of multi-threading, but we use our experience in application development to
support our claim that the threads behave correctly, and as a programmer would
expect. 
\subsection{Applications}
\subsubsection{Blinker (B)}
This application was built on top of the app-blinker-chain
example~\cite{blinker} and modified to remove the delays and LED activations, so
that it just passes data through three tasks until a maximum count is reached,
and then the thread terminates. This app was primarily used to demonstrate the
initial functionality of the system, and then to test mutexes. In both cases a
combination of \texttt{LOG} statements and final output values were used to
confirm the correct operation of the system. 
\subsubsection{Cuckoo-RSA (CRSA)}
This application is a combination of a cuckoo filter that inserts and queries 64
values a Bloom filter-like data structure~\cite{Chain}, and RSA encryption of message
using 128-bit keys~\cite{Chain}. Unfortunately this is the only application
combination where merging the original applications did not produce the same
final result as running the applications separately. Several bytes at the end of
a crucial array for RSA were always overwritten when the applications were
merged, and we were unable to fix the problem. Unit tests running the RSA
application in single threaded mode through our scheduler and using our
multi-threading library produce the correct result, but any combination of RSA
and cuckoo filtering produces the incorrect result. However, RSA does run to
completion, so we believe the results from C-RSA are still useful. 
\subsubsection{Cuckoo-Bitcount (CB)}
This application is a combination of the cuckoo filter described above and a
application that uses different algorithms to count the number of bits in an
array of integers~\cite{bitcount}. CB is a correctly working combination of two
applications that produces the same output as separate executions of the
unmodified versions of these applications. It expands the reliability of our
multi-threading library beyond simple applications like Blinker.
\subsubsection{Cuckoo-Bitcount-Blinker(CBB)}
CBB combines the cuckoo filter and bit count applications with two threads of
the blinker application. This results in four active threads at once, which is
the maximum number of threads our prototype implementation allows. We confirmed
that all four threads are fairly serviced by the round robin scheduler and they
correctly start and end.

>>>>

\usepackage{balance}
\usepackage{comment}
\usepackage{listings}
\usepackage{url}
\usepackage{xspace}
\usepackage[export]{adjustbox}
\usepackage{wrapfig}
\newcommand{\chain}{Chain\xspace}

\usepackage{color}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.95}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=none,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=4
    }
    \lstset{style=mystyle}

\numberofauthors{3}

@@

\section{Implementation}
\subsection{Scheduler}
\begin{figure*}
\begin{minipage}[b]{.48\textwidth}
\begin{lstlisting}[language=C]
    typedef struct thread_t {
        unsigned thread_id;
        context_t context;
    } thread_t;

    typedef struct thread_state_t {
        thread_t thread;
        unsigned active;
    } thread_state_t;

    // Array of running threads
    thread_state_t threads[MAX_NUM_THREADS];

    // List of free indicies in threads[]
    unsigned free_indicies[MAX_NUM_THREADS];
    unsigned free_indicies_size;

    // Index of current running thread in threads[]
    // Stored in self channel
    unsigned current;
\end{lstlisting}
\caption{Scheduler Data Structures}\label{schDataStruct}
\end{minipage}\hfill
\end{figure*}

The data structures for our scheduler are listed in figure \ref{schDataStruct}.
Like the rest of chain, memory for data structures is allocated at compile
time. Thus, the limit on the maximum number of threads allowed in a system is
fixed arbitrarily - not ideal, but there isn't a better option while operating
inside chain. The \texttt{threads} array is stored in all-to-one channel, from
every task to the scheduler task. The \texttt{free\_indicies} array is stored
in a one-to-all channel, from the scheduler task to every task.

Neither all-to-one nor one-to-all channels fit into standard \chain channel
semantics, although they resemble multicast-in and multicast-out channels
described in the original \chain paper. We eschew the notion that channels
have fixed start and end points and instead treat the channel as a general
location in non-volatile memory that can be read from and written to by
arbitrary tasks. However, accesses to these channels are encapsulated by our
library, so the programmer has no need to think about idempotency violations.

With these multicast channels, the scheduler task can act as a buffer between
multiple threads - arbitrary tasks write \texttt{threads[]} with calls to
\texttt{thread\_create} and \texttt{thread\_end} and tasks read from
\texttt{free\_indicies} through calls to thread create, maintaining
idempotency. Inside the scheduler task, we read and write \texttt{current},
so \texttt{current} is double buffered. The only issue is with
\texttt{thread\_init} since it is called from user-defined tasks and
initializes \texttt{current} - the same task that called \texttt{thread\_init}
could call \texttt{thread\_create}, and the \texttt{thread\_create} call
wouldn't see the update to the variable. Since \texttt{current} is always
initialized to the same value (0), we opt for manually swapping the buffer on
\texttt{current} so that other threads can see the update.

We keep an unsigned integer, \texttt{curr\_free\_index}, in volatile memory,
zeroed on every reboot, that acts as a current pointer in
\texttt{free\_indicies[]}. Every time the scheduler task runs,
\texttt{free\_indicies[]} is updated with the state of \texttt{threads[]},
based on whether or not the active bit is set for a given index. When
\texttt{thread\_create} is called, if \texttt{curr\_free\_index} is greater
than \texttt{free\_indicies\_size}, then \texttt{thread\_create} will fail in
POSIX fashion - the system does not have enough resources to support another
thread. Otherwise, \texttt{threads[free\_indicies[curr\_free\_index]]} is set
to the new thread. Upon power failure, \texttt{curr\_free\_index} is reset to
zero, so as long as the rest of the operations in the task are idempotent, so
will \texttt{thread\_create} calls - the same locations in \texttt{theads[]}
will be overwritten.

% TODO picture of array?

Mutexes (Simple, meant for pseudo channels) - Emily\\
Interrupts (mucked with IE flags) - Neil\\
% Interrupt prologue
% Interrupt setup
%\section{Implementation}
%Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
%Mutexes (Simple, meant for pseudo channels) - Emily\\
%Interrupts (mucked with IE flags) - Neil\\
%% Interrupt prologue
%% Interrupt setup

\section{Evaluation}
To test our multithreading library and mutex implementation, we instrumented
several existing Chain applications that do not share any tasks or channels to
run concurrently. The programs were run on either an msp430fr5969 launchpad or
on a Capybara sensing node, a custom intermittent development board with an
msp430fr5949 MCU. Reboot tests were carried out strictly on the Capybara board
because was simpler to guarantee the isolation of the MCU from the programming
fet than on the launchpad. We did not seekt to fully verify our implementation
of multi-threading, but we use our experience in application development to
support our claim that the threads behave correctly, and as a programmer would
expect.

\subsection{Applications}
\subsubsection{Blinker (B)}
This application was built on top of the app-blinker-chain
example~\cite{blinker} and modified to remove the delays and LED activations, so
that it just passes data through three tasks until a maximum count is reached,
and then the thread terminates. This app was primarily used to demonstrate the
initial functionality of the system, and then to test mutexes. In both cases a
combination of \texttt{LOG} statements and final output values were used to
confirm the correct operation of the system.
\subsubsection{Cuckoo-RSA (CRSA)}
This application is a combination of a cuckoo filter that inserts and queries 64
values a Bloom filter-like data structure~\cite{Chain}, and RSA encryption of message
using 128-bit keys~\cite{Chain}. Unfortunately this is the only application
combination where merging the original applications did not produce the same
final result as running the applications separately. Several bytes at the end of
a crucial array for RSA were always overwritten when the applications were
merged, and we were unable to fix the problem. Unit tests running the RSA
application in single threaded mode through our scheduler and using our
multi-threading library produce the correct result, but any combination of RSA
and cuckoo filtering produces the incorrect result. However, RSA does run to
completion, so we believe the results from C-RSA are still useful.
\subsubsection{Cuckoo-Bitcount (CB)}
This application is a combination of the cuckoo filter described above and a
application that uses different algorithms to count the number of bits in an
array of integers~\cite{bitcount}. CB is a correctly working combination of two
applications that produces the same output as separate executions of the
unmodified versions of these applications. It expands the reliability of our
multi-threading library beyond simple applications like Blinker.
\subsubsection{Cuckoo-Bitcount-Blinker(CBB)}
CBB combines the cuckoo filter and bit count applications with two threads of
the blinker application. This results in four active threads at once, which is
the maximum number of threads our prototype implementation allows. We confirmed
that all four threads are fairly serviced by the round robin scheduler and they
correctly start and end.
